SSwAB-IT Phase 1 Analysis

### 2024-04-04 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
Update phase_1_analysis.qmd

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Notes/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
editor: source
---

@@

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Format
df <- df %>%
  select(
    record_id, redcap_data_access_group,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
# # Drop levels
# level_drop <- c("sex", "Race", "Ethnicity")
# db_demographics[, level_drop] <- lapply(
#   db_demographics[, level_drop],
#   function(c) {
#     droplevels(factor(c))
#   }
# )
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Bug list
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
# Convert yes/no to positive/negative
df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Positive")
})
df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Positive")
})
df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
# Add overall to list
bugs <- c("overall", bugs)
bug_names <- c("Across all bugs", bug_names)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Agreement columns
df[, bugs] <- lapply(bugs, function(b) {
  cos <- df[, paste0(b, "_cos")]
  sos <- df[, paste0(b, "_sos")]
  f <- factor(paste(cos, sos, sep = "_"),
    exclude = c(
      "NA_Negative", "Negative_NA", "NA_Positive", "Positive_NA", "NA_NA"
    )
  )
  f <- factor(f,
    levels = c(
      "Negative_Negative", "Negative_Positive", "Positive_Negative",
      "Positive_Positive"
    ),
    labels = c(
      "COS Negative / SOS Negative", "COS Negative / SOS Positive",
      "COS Positive / SOS Negative", "COS Positive / SOS Positive"
    )
  )
  return(f)
})
```

@@

```{r}
db_demographics %>%
  select(group, sex, Race, Ethnicity) %>%
  tbl_summary(by = group, label = list("sex" = "Sex")) %>%
  add_overall() %>%
  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
  separate_p_footnotes() %>%
  as_gt()
```

>>>>

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

@@

```{r data cleaning}
#| include: false
# Import from REDCap
unlockREDCap(c(rcon = "SSwAB-IT (New)"),
  keyring = "API_KEYs",
  envir = 1,
  url = "https://redcap.ucdenver.edu/api/"
)
exportBulkRecords(list(db = rcon), envir = 1)
# Exclude
exclude <- c("01-030_W", "01-038_W", "02-0038", "02-0039", "04-001")
# Combine collection results
df <- full_join(db_cos_collection_results, db_sos_collection_results)
# Add dates to demographics
db_demographics <- left_join(db_demographics,
  df %>% select(record_id, cos_result_date),
  by = join_by(record_id)
)
db_demographics$age_at_cos <- difftime(db_demographics$cos_result_date,
  db_demographics$dob,
  units = "days"
) / 365.25
# Format
df <- df %>%
  select(
    record_id, redcap_data_access_group,
    staphylococcus_aureus_cos:other_result_cos,
    staphylococcus_aureus_sos:other_result_sos
  ) %>%
  filter(!record_id %in% exclude)
# Remove empty demographics rows
db_demographics <- db_demographics %>%
  filter(
    redcap_event_name == "Clinic Visit (Arm 1: Eligible)",
    !record_id %in% exclude
  )
# Excluded participants
db_medications <- db_medications %>% filter(!record_id %in% exclude)
# Combine race and ethnicity columns
races <- c(
  "American Indian or Alaskan Native", "Asian", "Hawaiian or Pacific Islander",
  "Black or African American", "White", "Unknown", "Other"
)
ethnicities <- c("Hispanic", "Non-Hispanic", "Unknown/Not Reported")
db_demographics$Race <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("race", names(r))] == "Checked")
  if (length(w) == 1) {
    return(races[w])
  } else if (length(w) > 1) {
    return("Multiple Races")
  } else if (length(w) == 0) {
    return(NA)
  }
})
db_demographics$Ethnicity <- apply(db_demographics, 1, function(r) {
  w <- which(r[grep("ethnicity", names(r))] == "Checked")
  if (length(w) == 1) {
    return(ethnicities[w])
  } else if (length(w) > 1) {
    return(paste(ethnicities[w], sep = "/"))
  } else if (length(w) == 0) {
    return(NA)
  }
})
# # Drop levels
# level_drop <- c("sex", "Race", "Ethnicity")
# db_demographics[, level_drop] <- lapply(
#   db_demographics[, level_drop],
#   function(c) {
#     droplevels(factor(c))
#   }
# )
# Add groups to other dataframes
db_medications <- left_join(db_medications,
  db_demographics %>% select(record_id, group),
  by = join_by(record_id)
)
# Bug list
bugs <- c(
  "staphylococcus_aureus", "haemophilus_influenza", "pseudomonas_aeruginosa",
  "pseudomonas_non_aero", "s_maltophilia", "burkholderia", "aspergillus", "ntm",
  "other"
)
bug_names <- c(
  "S. aureus", "H. influenza", "P. aeruginosa", "Pseudomonas non-aeruginosa",
  "S. maltophilia", "Burkholderia", "Aspergillus", "NTM", "Other"
)
# Convert yes/no to positive/negative
df[, paste0(bugs, "_cos")] <- lapply(df[, paste0(bugs, "_cos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
df[, paste0(bugs, "_sos")] <- lapply(df[, paste0(bugs, "_sos")], factor,
  levels = c("No", "Yes"), labels = c("Negative", "Positive")
)
# Create overall columns (any bug detected in sample)
df$overall_cos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_cos")] == "Positive")
})
df$overall_cos <- factor(df$overall_cos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
df$overall_sos <- apply(df, 1, function(r) {
  any(r[paste0(bugs, "_sos")] == "Positive")
})
df$overall_sos <- factor(df$overall_sos,
  levels = c(F, T), labels = c("Negative", "Positive")
)
# Add overall to list
bugs <- c("overall", bugs)
bug_names <- c("Across all bugs", bug_names)
bug_list <- as.list(bug_names)
names(bug_list) <- bugs
# Agreement columns
df[, bugs] <- lapply(bugs, function(b) {
  cos <- df[, paste0(b, "_cos")]
  sos <- df[, paste0(b, "_sos")]
  f <- factor(paste(cos, sos, sep = "_"),
    exclude = c(
      "NA_Negative", "Negative_NA", "NA_Positive", "Positive_NA", "NA_NA"
    )
  )
  f <- factor(f,
    levels = c(
      "Negative_Negative", "Negative_Positive", "Positive_Negative",
      "Positive_Positive"
    ),
    labels = c(
      "COS Negative / SOS Negative", "COS Negative / SOS Positive",
      "COS Positive / SOS Negative", "COS Positive / SOS Positive"
    )
  )
  return(f)
})
```

@@

```{r}
db_demographics %>%
  select(group, age_at_cos, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = group,
    statistic = all_continuous()~c("{median} ({min}, {max})"),
    label = list("sex" = "Sex", "age_at_cos" = "Age at COS")
  ) %>%
  add_overall() %>%
  add_p(test = list(c("Race", "Ethnicity") ~ "fisher.test")) %>%
  separate_p_footnotes() %>%
  as_gt()
```
